---
title: "Food Deserts in the United States"
author: "Team B"
date: "7/28/2020"
output: 
  html_document:
   code_folding: hide
   number_sections: true
   toc: yes
   toc_depth: 3
   toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```


```{r xkablesummary}
loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")
loadPkg("leaps")
loadPkg("ggplot2")
loadPkg("ggmap")
loadPkg("ISLR")
loadPkg("faraway")
loadPkg("tigris")
loadPkg("acs")
loadPkg("stringr") 


xkabledply = function(modelsmmrytable, title="Table", digits = 4, pos="left", bso="striped", wide=FALSE) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display model summary. 
  #' wrapper for the base::summary function on model objects
  #' Can also use as head for better display
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param modelsmmrytable This can be a generic table, a model object such as lm(), or the summary of a model object summary(lm()) 
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return HTML table for display
  #' @examples
  #' library("xtable")
  #' library("kableExtra")
  #' xkabledply( df, title="Table testing", pos="left", bso="hover" )
  #' xkabledply( ISLR::Hitters[1:5,] )
  if (wide) { modelsmmrytable <- t(modelsmmrytable) }
  modelsmmrytable %>%
    xtable() %>% 
    kable(caption = title, digits = digits) %>%
    kable_styling(bootstrap_options = bso, full_width = FALSE, position = pos)
}

xkabledplyhead = function(df, rows=5, title="Head", digits = 4, pos="left", bso="striped") { 
  xkabledply(df[1:rows, ], title, digits, pos, bso, wide=FALSE)
}

xkabledplytail = function(df, rows=5, title="Tail", digits = 4, pos="left", bso="striped") { 
  trows = nrow(df)
  xkabledply(df[ (trows-rows+1) : trows, ], title, digits, pos, bso, wide=FALSE)
}

xkablesummary = function(df, title="Table: Statistics summary.", digits = 4, pos="left", bso="striped") { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param df The dataframe.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return The HTML summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters, title="Five number summary", pos="left", bso="hover"  )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  xkabledply(s, title=title, digits = digits, pos=pos, bso=bso )
}

xkablevif = function(model, title="VIFs of the model", digits = 3, pos="left", bso="striped", wide=TRUE) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param model The lm or compatible model object.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return The HTML summary table of the VIFs for a model for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( lm(Salary~Hits+RBI, data=ISLR::Hitters, wide=T ) )
  
  vifs = table( names(model$coefficients)[2:length(model$coefficients)] ) # remove intercept to set column names
  vifs[] = faraway::vif(model) # set the values
  if (wide) { vifs <- t(vifs) }
  xkabledply( vifs, title=title, digits = digits, pos=pos, bso=bso )
}
```


# Chapter 1 Introduction 
For years social scientists have used various parameters to define food deserts. Food dese

## Exploratory Data Analysis

```{r Using Data Frame and Picking Variables}
### now we use dataset food on the master branch which has been extracted 
food <- data.frame(read.csv("food.csv"))
dim(food)
food$State<-as.factor(food$State)
food$Region <- as.factor(food$Region)


# Changed y variable name 
colnames(food)[8] <- c("Desert")
#str(food)

# Deciding Best Varaible with Forward Selection and Subsetting
reg.10.forward <-regsubsets(Desert~.-CensusTract-State-TractLOWI-TractKids-TractSeniors, data = food, nvmax=10, nbest=1, method="forward")
plot(reg.10.forward, scale = "adjr2", main = "Adjusted R^2")
plot(reg.10.forward, scale = "Cp", main = "CP")
plot(reg.10.forward, scale = "bic", main = "BIC")

# Backward Selection
reg.best10.bkwd <- regsubsets(Desert~.-CensusTract-State-TractLOWI-TractKids-TractSeniors, data= food, nvmax=10, nbest=1, method="backward")
plot(reg.best10.bkwd, scale = "adjr2", main = "Adjusted R^2")
plot(reg.best10.bkwd, scale = "bic", main = "BIC")
plot(reg.best10.bkwd, scale = "Cp", main = "CP")

# Sequential Selection 
reg.best10.seq <- regsubsets(Desert~CensusTract+State+TractLOWI+TractKids+TractSeniors, data= food, nvmax=10, nbest=1, method="seqrep")
plot(reg.best10.seq, scale = "adjr2", main = "Adjusted R^2")
plot(reg.best10.seq, scale = "Cp", main = "CP")
plot(reg.best10.seq, scale = "bic", main = "BIC")

#Correlation Tests on x variables and y 
cor.test(food$POP2010, food$Desert)
cor.test(food$TractBlack, food$Desert)
cor.test(food$PovertyRate, food$Desert)
cor.test(food$TractSNAP, food$Desert)
```





```{r}
### sampling process removed, new "food.csv" has been uploaded
### Model Building
summary(lm(Desert~.-CensusTract-State,data = food))
glm_1sttry<-glm(Desert~.-CensusTract-State,data = food,family = "binomial")
summary(glm_1sttry)
exp(glm_1sttry$coefficients)

```




```{r}
#Eric ANOVA building test:
summary(food)

my_data <- food

set.seed(123)
dplyr::sample_n(my_data,10)

levels(my_data$State)

#Creating levels as seen under the my_data

my_data$State <- ordered(my_data$POP2010) 
                         levels = c("State", "Urban", "POP2010","OHU2010", "NUMGQTRS", "LILATracts_1And10]", "PovertyRate", "MedianFamilyIncome",
                                    "TractLOWI", "TractKids", "TractSenios", "TractWhite", "TractBlack", "TractAsian", "TractNHOPI", "TractAIAN", "TractOMultir", 
                                    "TractHispanic","TractHUNV","TractSNAP")
library(dplyr)
group_by(my_data, POP2010) %>%
  summarise(
        count = n(),
        mean = mean(Urban, na.rm = TRUE),
        sd = sd(Urban, na.rm = TRUE)
                           )
install.packages("ggpubr")


#Used this as a test, got a plot with some data points.. Not sure how to get more on the Y axis? (Not numeric here perhaps?)

library("ggpubr")
ggboxplot(my_data, x = "Urban", y = "POP2010", 
          color = "POP2010", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("ctrl", "trt1", "trt2"),
          ylab = "POP2010", xlab = "Urban")



# Box plot
boxplot(POP2010 ~ Urban, data = my_data,
        xlab = "Urban", ylab = "POP2010",
        frame = FALSE, col = c("#00AFBB", "#E7B800", "#FC4E07"))


#More Plots

library("gplots")
mean(POP2010 ~ Urban, data = my_data, frame = FALSE,
          xlab = "Urban", ylab = "POP2010",
          main="Mean Plot with 95% CI") 


# Compute the analysis of variance
res.aov <- aov(POP2010 ~ Urban, data = my_data)

# Summary of the analysis

summary(res.aov)

####

```

```{r}
# Ashley Chi-Square (only works with two categotical variables)
chisq.test(food$Urban,y=food$Desert)
chisq.test(food$Region, y=food$Desert)
chisq.test(food$State, y=food$Desert)
```



```{r}
#Eric ANOVA building test:

summary(food)

my_data <- food

set.seed(123)
dplyr::sample_n(my_data,10)

levels(my_data$State)

#Creating levels as seen under the my_data

my_data$State <- ordered(my_data$POP2010) 
                         levels = c("State", "Urban", "POP2010","OHU2010", "NUMGQTRS", "LILATracts_1And10", "PovertyRate", "MedianFamilyIncome",
                                    "TractLOWI", "TractKids", "TractSenios", "TractWhite", "TractBlack", "TractAsian", "TractNHOPI", "TractAIAN", "TractOMultir", 
                                    "TractHispanic","TractHUNV","TractSNAP")
library(dplyr)
group_by(my_data, POP2010) %>%
  summarise(
        count = n(),
        mean = mean(Urban, na.rm = TRUE),
        sd = sd(Urban, na.rm = TRUE)
                           )
install.packages("ggpubr")


#Used this as a test, got a plot with some data points.. Not sure how to get more on the Y axis? (Not numeric here perhaps?)

library("ggpubr")
ggboxplot(my_data, x = "Urban", y = "POP2010", 
          color = "POP2010", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("ctrl", "trt1", "trt2"),
          ylab = "POP2010", xlab = "Urban")



# Box plot
boxplot(POP2010 ~ Urban, data = my_data,
        xlab = "Urban", ylab = "POP2010",
        frame = FALSE, col = c("#00AFBB", "#E7B800", "#FC4E07"))


#More Plots

library("gplots")
mean(POP2010 ~ Urban, data = my_data, frame = FALSE,
          xlab = "Urban", ylab = "POP2010",
          main="Mean Plot with 95% CI") 


# Compute the analysis of variance
res.aov <- aov(POP2010 ~ Urban, data = my_data)

# Summary of the analysis

summary(res.aov)
```



```{r}
#### Di Anova copy and paste part
res.aov_1 <- aov(POP2010 ~ Urban, data = my_data)
res.aov_2 <- aov(POP2010 ~ State, data = my_data)
res.aov_3 <- aov(POP2010 ~ Region, data = my_data)
res.aov_4 <- aov(POP2010 ~ OHU2010, data = my_data)
res.aov_5 <- aov(POP2010 ~ NUMGQTRS, data = my_data)
res.aov_6 <- aov(POP2010 ~ LILATracts_1And10, data = my_data)
res.aov_7 <- aov(POP2010 ~ PovertyRate, data = my_data)
res.aov_8 <- aov(POP2010 ~ MedianFamilyIncome, data = my_data)
res.aov_9 <- aov(POP2010 ~ TractLOWI, data = my_data)
res.aov_10 <- aov(POP2010 ~ TractKids, data = my_data)
res.aov_11 <- aov(POP2010 ~ TractSeniors, data = my_data)
res.aov_12 <- aov(POP2010 ~ TractWhite, data = my_data)
res.aov_13 <- aov(POP2010 ~ TractBlack, data = my_data)
res.aov_14 <- aov(POP2010 ~ TractAsian, data = my_data)
res.aov_15 <- aov(POP2010 ~ TractNHOPI, data = my_data)
res.aov_16 <- aov(POP2010 ~ TractAIAN, data = my_data)
res.aov_17 <- aov(POP2010 ~ TractOMultir, data = my_data)
res.aov_18 <- aov(POP2010 ~ TractHispanic, data = my_data)
res.aov_19 <- aov(POP2010 ~ TractHUNV, data = my_data)
res.aov_20 <- aov(POP2010 ~ TractSNAP, data = my_data)
```



res.aov_1 <- aov(Desert ~ Urban, data = my_data)
res.aov_2 <- aov(Desert ~ State, data = my_data)
res.aov_3 <- aov(Desert ~ Region, data = my_data)
res.aov_4 <- aov(Desert ~ OHU2010, data = my_data)
res.aov_5 <- aov(Desert ~ NUMGQTRS, data = my_data)
res.aov_6 <- aov(Desert ~ POP2010, data = my_data)
res.aov_7 <- aov(Desert ~ PovertyRate, data = my_data)
res.aov_8 <- aov(Desert ~ MedianFamilyIncome, data = my_data)
res.aov_9 <- aov(Desert ~ TractLOWI, data = my_data)
res.aov_10 <- aov(Desert ~ TractKids, data = my_data)
res.aov_11 <- aov(Desert ~ TractSenios, data = my_data)
res.aov_12 <- aov(Desert ~ TractWhite, data = my_data)
res.aov_13 <- aov(Desert ~ TractBlack, data = my_data)
res.aov_14 <- aov(Desert ~ TractAsian, data = my_data)
res.aov_15 <- aov(Desert ~ TractNHOPI, data = my_data)
res.aov_16 <- aov(Desert ~ TractAIAN, data = my_data)
res.aov_17 <- aov(Desert ~ TractOMultir, data = my_data)
res.aov_18 <- aov(Desert ~ TractHispanic, data = my_data)
res.aov_19 <- aov(Desert ~ TractHUNV, data = my_data)
res.aov_20 <- aov(Desert ~ TractSNAP, data = my_data)









